Fast Hyper-Parameter Selection for Graph Kernels via Subsampling and Multiple Kernel Learning

### 2016-11-25 paper.tex
Modified Lambda of EasyMKL

EasyMKL finds the coefficients $\boldsymbol{\eta}$ that maximize the margin on the training set, where the margin is computed as the distance between the convex hulls of positive and negative examples. In particular, the general problem EasyMKL tries to optimize is the following:
\vspace{-0.4cm}
\begin{equation}
\label{eq:easymklotig}
\max_{\boldsymbol{\eta}:||\boldsymbol{\eta}||_2=1} \min_{\boldsymbol{\gamma} \in \Gamma} \mbox { }  \boldsymbol{\gamma}^{\top} \YY ( \sum_{r=0}^R \eta_r \KK_r) \YY \boldsymbol{\gamma} + \lambda ||\boldsymbol{\gamma}||^2.
\end{equation}
\vspace{-0.3cm}

where $\YY$ is a diagonal matrix with training labels on the diagonal, and $\lambda$ is a regularization hyper-parameter. The domain $\Gamma$ represents two probability distributions over the set of positive and negative examples of the training set, that is $\Gamma = \{\boldsymbol{\gamma} \in \Real_+^{\ell} \hspace{0.2cm} |  \sum_{y_i = +1} \boldsymbol{\gamma}_i = 1 , \sum_{y_i = -1} \boldsymbol{\gamma}_i = 1\}.
$
Note that any element $\boldsymbol{\gamma} \in \Gamma$ corresponds to a pair of points, the first in the convex hull of positive training examples and the second in the convex hull of negative training examples.
At the solution, the first term of the objective function represents the obtained margin, that is the (squared) distance between a point in the convex hull of positive examples and a point in the convex hull of negative examples, in the compounded feature space.

@@

\section{Experimental Assessment}
We conducted our experiments on five real-world datasets (see \cite{DaSanMartino2016} for a more detailed description).
%AIDS~\cite{Weislow1989a}, CAS\footnote{http://cheminformatics.org/datasets/bursi/},
%CPDB~\cite{Helma2004a}, NCI1~\cite{Wale2006} and GDD~\cite{dobson2003}.
%The first four datasets contain chemical particles encoded in graph form, while the latter encodes proteins.
All the nodes are labeled and there are no self loops.
The AIDS dataset contains chemical compounds (encoded in graph form), labeled according to
their antiviral activity; CAS and CPDB store compounds labeled according to their mutagenicity; NCI1 consists of chemical compounds screened for activity against non-small lung cancer cells; GDD is composed of X-ray crystal structures of
proteins.
All the datasets represent binary classification problems.

@@

To learn $\hat{\boldsymbol{\eta}}$ we chose a random subsampling of training data and perform a 10-fold cross validation using EasyMKL to combine kernels and SVM as base learner. Then we fit the model with the best combination and select the top kernels.
In the second step we calculate the top kernel matrices using the whole training data and combine it to get the kernel. We propose 3 different methods to perform this combination:
(1)\ksmkl: using the EasyMKL algorithm, (2)\kssum: calculating the average of kernels, (3)\ksw: using the same weights (rescaled) from the first learning step.\\
For testing, we compared our method against 3 hard baselines, that are: (a)\kval: the kernel is selected in the nested-cv procedure, (b)\ksum: the kernel is the average of all base kernels, (c)\kmkl: we use EasyMKL to combine the whole set of kernels.
In all models, we use a SVM as base learner, and we select the hyperparameters using a nested 10-fold CV procedure. The hyperparameters are $\lambda\in\{0,0.1\dots0.4,0.6\dots1\}$ for models where we use EasyMKL, and $C\in\{2^i, i:-3\dots6\}$.
Note that for all baselines we need to calculate all kernel matrices using all of training data directly.\\
The kernels used in our experiments are: all NSPDK combinations with $h \in \{0,1\dots9\}$ and $d \in \{0,1\dots9\}$, all ODDST combinations with $h \in \{0,1\dots9\}$ and $\lambda \in \{0.5, 0.7,0.8,0.9,1.0,1.1,1.2,1.4,1.6,1.8\}$, all WL with $h \in \{0,1\dots9\}$.

>>>>

EasyMKL finds the coefficients $\boldsymbol{\eta}$ that maximize the margin on the training set, where the margin is computed as the distance between the convex hulls of positive and negative examples. In particular, the general problem EasyMKL tries to optimize is the following:
\vspace{-0.4cm}
\begin{equation}
\label{eq:easymklotig}
\max_{\boldsymbol{\eta}:||\boldsymbol{\eta}||_2=1} \min_{\boldsymbol{\gamma} \in \Gamma} \mbox { }  \boldsymbol{\gamma}^{\top} \YY ( \sum_{r=0}^R \eta_r \KK_r) \YY \boldsymbol{\gamma} + \Lambda ||\boldsymbol{\gamma}||^2.
\end{equation}
\vspace{-0.3cm}

where $\YY$ is a diagonal matrix with training labels on the diagonal, and $\Lambda$ is a regularization hyper-parameter. The domain $\Gamma$ represents two probability distributions over the set of positive and negative examples of the training set, that is $\Gamma = \{\boldsymbol{\gamma} \in \Real_+^{\ell} \hspace{0.2cm} |  \sum_{y_i = +1} \boldsymbol{\gamma}_i = 1 , \sum_{y_i = -1} \boldsymbol{\gamma}_i = 1\}.
$
Note that any element $\boldsymbol{\gamma} \in \Gamma$ corresponds to a pair of points, the first in the convex hull of positive training examples and the second in the convex hull of negative training examples.
At the solution, the first term of the objective function represents the obtained margin, that is the (squared) distance between a point in the convex hull of positive examples and a point in the convex hull of negative examples, in the compounded feature space.

@@

\section{Experimental Assessment}
We conducted our experiments on five real-world datasets (see \cite{DaSanMartino2016} for a more detailed description).
%AIDS~\cite{Weislow1989a}, CAS\footnote{http://cheminformatics.org/datasets/bursi/},
%CPDB~\cite{Helma2004a}, NCI1~\cite{Wale2006} and GDD~\cite{dobson2003}.
%The first four datasets contain chemical particles encoded in graph form, while the latter encodes proteins.
All the nodes are labeled and there are no self loops.
The AIDS dataset contains chemical compounds (encoded in graph form), labeled according to
their antiviral activity; CAS and CPDB store compounds labeled according to their mutagenicity; NCI1 and NCI109 consist of chemical compounds screened for activity against non-small lung cancer cells; GDD is composed of X-ray crystal structures of
proteins.
All the datasets represent binary classification problems.

@@

To learn $\hat{\boldsymbol{\eta}}$ we chose a random subsampling of training data and perform a 10-fold cross validation using EasyMKL to combine kernels and SVM as base learner. Then we fit the model with the best combination and select the top kernels.
In the second step we calculate the top kernel matrices using the whole training data and combine it to get the kernel. We propose 3 different methods to perform this combination:
(1)\ksmkl: using the EasyMKL algorithm, (2)\kssum: calculating the average of kernels, (3)\ksw: using the same weights (rescaled) from the first learning step.\\
For testing, we compared our method against 3 hard baselines, that are: (a)\kval: the kernel is selected in the nested-cv procedure, (b)\ksum: the kernel is the average of all base kernels, (c)\kmkl: we use EasyMKL to combine the whole set of kernels.
In all models, we use a SVM as base learner, and we select the hyperparameters using a nested 10-fold CV procedure. The hyperparameters are $\Lambda\in\{0,0.1\dots0.4,0.6\dots1\}$ for models where we use EasyMKL, and $C\in\{2^i, i:-3\dots6\}$.
Note that for all baselines we need to calculate all kernel matrices using all of training data directly.\\
The kernels used in our experiments are: all NSPDK combinations with $h \in \{0,1\dots9\}$ and $d \in \{0,1\dots9\}$, all ODDST combinations with $h \in \{0,1\dots9\}$ and $\lambda \in \{0.5, 0.7,0.8,0.9,1.0,1.1,1.2,1.4,1.6,1.8\}$, all WL with $h \in \{0,1\dots9\}$.
